---
title: "Mtcars_DataAnalysis"
author: "Himanshi"
date: "2025-08-08"
output:
  pdf_document: default
  html_document: default
---

## Project Overview
This project analyzes the `mtcars` dataset, which contains performance and design specifications for 32 car models from 1974 Motor Trend magazine.  
The goal is to explore relationships between variables, compare groups, and build a simple regression model to understand what affects fuel efficiency (miles per gallon, `mpg`).

---

## Loading the Data
```{r}
data(mtcars)

# Basic overview
head(mtcars)
summary(mtcars)
```
# Regression
```{r}
model <- lm(mpg ~ wt + hp, data=mtcars)
summary(model)
plot(mtcars$wt, mtcars$mpg, pch=19)
abline(lm(mpg ~ wt, data=mtcars), col="red")
```

# Correlation Heatmap
```{r}
install.packages("corrplot", repos = "https://cloud.r-project.org")
library(corrplot)
corrplot(cor(mtcars), method="color", type="upper", tl.col="black", tl.srt=45)
```

# Pairwise Scatter Plots
```{r}
pairs(~ mpg + hp + wt + disp, data=mtcars,
      main="Pairwise Plots of Selected Variables")
```

# BoxPlots by Category
```{r}
boxplot(mpg ~ am, data=mtcars,
        names=c("Automatic", "Manual"),
        col=c("orange", "lightblue"),
        main="MPG by Transmission Type")
```

# Multiple linear regression
```{r}
model <- lm(mpg ~ wt + hp, data=mtcars)
summary(model)
```

# Visualizing fit for weight
```{r}
plot(mtcars$wt, mtcars$mpg, pch=19, col="blue")
abline(lm(mpg ~ wt, data=mtcars), col="red", lwd=2)
```

# Residual diagnostics
```{r}
par(mfrow=c(2,2))
plot(model)
```