---
title: "PlantGrowth_ExperimentalDesign"
author: "Himanshi"
date: "2025-08-08"
output: html_document
---

## Introduction
The **PlantGrowth** dataset contains results from an experiment comparing the effect of different treatments on plant weight.  
There are three groups:
- **ctrl**: Control group  
- **trt1**: Treatment 1  
- **trt2**: Treatment 2  

The dataset has 30 observations.

data(PlantGrowth)
boxplot(weight ~ group, data=PlantGrowth, main="Plant Growth by Group")

# View dataset
```{r}
head(PlantGrowth)
summary(PlantGrowth)
table(PlantGrowth$group)
```

# Descriptive Statistics
```{r}
install.packages("dplyr", repos = "https://cloud.r-project.org")
library(dplyr)
install.packages("ggplot2", repos = "https://cloud.r-project.org")
library(ggplot2)

PlantGrowth %>%
  group_by(group) %>%
  summarise(
    mean_weight = mean(weight),
    sd_weight = sd(weight),
    n = n()
  )
```

# Boxplot with jittered Points
```{r}
install.packages("ggplot2", repos = "https://cloud.r-project.org")
library(ggplot2)

ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_boxplot(alpha = 0.6) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.7) +
  labs(title = "Plant Weight by Group",
       x = "Group",
       y = "Weight") +
  theme_minimal()
```


# Mean Plot with Error Bars
```{r}
install.packages("dplyr", repos = "https://cloud.r-project.org")
library(ggplot2)
install.packages("ggplot2", repos = "https://cloud.r-project.org")
library(dplyr)

stats <- PlantGrowth %>%
  group_by(group) %>%
  summarise(
    mean_weight = mean(weight),
    se = sd(weight) / sqrt(n())
  )
ggplot(stats, aes(x = group, y = mean_weight, fill = group)) +
  geom_col(alpha = 0.7) +
  geom_errorbar(aes(ymin = mean_weight - se, ymax = mean_weight + se),
                width = 0.2, color = "black") +
  geom_text(aes(label = round(mean_weight, 2)), vjust = -1) +
  labs(title = "Mean Plant Weight by Group",
       x = "Group",
       y = "Mean Weight") +
  theme_minimal()
```

# ANOVA
```{r}
model <- aov(weight ~ group, data=PlantGrowth)
summary(model)
```

# Post-hoc test
```{r}
TukeyHSD(model)
```

# Violin Plot
```{r}
ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, fill = "white") +
  labs(title = "Distribution of Plant Weights by Group",
       x = "Group",
       y = "Weight") +
  theme_minimal()

```