---
title: "Titanic_Survival_Analysis"
author: "Himanshi"
date: "2025-08-08"
output: html_document
---

## Project Overview
This project analyzes the `Titanic` dataset from R's built-in `datasets` package.  
It contains aggregated passenger data by Class, Sex, Age, and Survival status.  
The goal is to explore survival patterns and understand how class, gender, and age affected outcomes.

---

## Loading and Inspecting the Data
```{r}
data(Titanic)
titanic_df <- as.data.frame(Titanic)

head(titanic_df)
str(titanic_df)
```

# Survival rate
```{r}
aggregate(Freq ~ Sex + Class + Survived, data=titanic_df, sum)
```

# Overall Survival rate
```{r}
total <- sum(titanic_df$Freq)
survived <- sum(titanic_df$Freq[titanic_df$Survived == "Yes"])
survival_rate <- survived / total * 100
survival_rate
```

# Survival By Sex
```{r}
sex_survival <- aggregate(Freq ~ Sex + Survived, data=titanic_df, sum)

barplot(
  height = matrix(sex_survival$Freq, nrow=2, byrow=TRUE),
  beside = TRUE,
  names.arg = c("Female", "Male"),
  col = c("green", "red"),
  legend = c("Survived", "Died"),
  main = "Survival by Sex",
  ylab = "Number of Passengers"
)
```


# Survival By Class
```{r}
class_survival <- aggregate(Freq ~ Class + Survived, data=titanic_df, sum)

barplot(
  height = matrix(class_survival$Freq, nrow=2, byrow=TRUE),
  beside = TRUE,
  names.arg = c("1st", "2nd", "3rd", "Crew"),
  col = c("green", "red"),
  legend = c("Survived", "Died"),
  main = "Survival by Class",
  ylab = "Number of Passengers"
)
```

# Heatmap: Class vs Sex vs Survival
```{r}
class_survival <- aggregate(Freq ~ Class + Survived, data=titanic_df, sum)

barplot(
  height = matrix(class_survival$Freq, nrow=2, byrow=TRUE),
  beside = TRUE,
  names.arg = c("1st", "2nd", "3rd", "Crew"),
  col = c("green", "red"),
  legend = c("Survived", "Died"),
  main = "Survival by Class",
  ylab = "Number of Passengers"
)
```